Put Together Simulations

### 2023-10-16 manuscript.Rmd
all the updates

---
title             : "Visualizing and Interpreting Multi-Group Confirmatory Factor Analysis"
shorttitle        : "VISUAL MGCFA"

@@

abstract: |
  Latent variable modeling as a lens for psychometric theory is a popular tool for social scientists to examine measurement of constructs [@beaujean2014]. Journals such as *Assessment* regularly publish articles supporting measures of latent constructs wherein a measurement model is established. Confirmatory factor analysis can be used to investigate the replicability and generalizability of the measurement model in new samples, while multi-group confirmatory factor analysis is used to examine the measurement model across groups within samples [@brown2015]. With the rise of the replication crisis and "psychology's renaissance" [@nelson2018], interest in divergence in measurement has increased, often focused on small parameter differences within the latent model. This manuscript outlines ways to visualize potential non-invariance, to supplement large numbers of tables that often overwhelm a reader within these published reports. Readers will learn how to interpret the impact and size of the proposed non-invariance in models. While it is tempting to suggest that problems with replication and generalizability are simply issues with measurement, it is crucial to remember that all models have variability and error, even those models estimating the differences between item functioning, such as multi-group confirmatory factor analysis.

@@

classoption       : "man"
bibliography: references.bib

@@

csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa7"
output            : papaja::apa6_pdf
---

```{r libraries, include = FALSE}
library(papaja)
library(lavaan)
library(broom)
library(dplyr)
library(ggplot2)
# devtools::install_github("psyteachr/introdataviz")
# devtools::install_github("ddueber/dmacs")
library(introdataviz)
library(dmacs)
library(cowplot)

@@

```{r cfa-function, include = FALSE}
# CFA function
mgcfa <- function(data, group, model){

  group_names <- unique(data[ , group])
  data$group <- data[ , group]

  model.overall <- cfa(model = model, data = data, 
                       meanstructure = T)
  model.group1 <- cfa(model = model, 
                      data = subset(data, group == group_names[1]),
                      meanstructure = T)
  model.group2 <- cfa(model = model, 
                      data = subset(data, group == group_names[2]),
                      meanstructure = T)
  model.configural <- cfa(model = model, data = data, 
                          group = group, meanstructure = T)
  model.metric <- cfa(model = model, data = data, 
                      group = group, meanstructure = T,
                      group.equal = "loadings")
  model.scalar <- cfa(model = model, data = data, 
                      group = group, meanstructure = T, 
                      group.equal = c("loadings", "intercepts"))
  model.strict <- cfa(model = model, data = data, 
                      group = group, meanstructure = T, 
                      group.equal =  c("loadings", "intercepts", "residuals"))

  model_coef <- bind_rows(
    tidy(model.overall, conf.level = .95) %>% mutate(model = "Overall"), 
    tidy(model.group1, conf.level = .95) %>% mutate(model = "Group 1"), 
    tidy(model.group2, conf.level = .95) %>% mutate(model = "Group 2"), 
    tidy(model.configural, conf.level = .95) %>% mutate(model = "Configural"), 
    tidy(model.metric, conf.level = .95) %>% mutate(model = "Metric"), 
    tidy(model.scalar, conf.level = .95) %>% mutate(model = "Scalar"), 
    tidy(model.strict, conf.level = .95) %>% mutate(model = "Strict")
  )

  model_fit <- bind_rows(
    glance(model.overall) %>% mutate(model = "Overall"), 
    glance(model.group1) %>% mutate(model = "Group 1"), 
    glance(model.group2) %>% mutate(model = "Group 2"), 
    glance(model.configural) %>% mutate(model = "Configural"), 
    glance(model.metric) %>% mutate(model = "Metric"), 
    glance(model.scalar) %>% mutate(model = "Scalar"), 
    glance(model.strict) %>% mutate(model = "Strict")
    )

  return(list(
    "model_coef" = model_coef,
    "model_fit" = model_fit,
    "model.overall" = model.overall,
    "model.group1" = model.group1,
    "model.group2" = model.group2,
    "model.configural" = model.configural,
    "model.metric" = model.metric,
    "model.scalar" = model.scalar,
    "model.strict" = model.strict
  ))

}
```

```{r test-function, eval = F, include = F}
# temp test
data_coef = results.large.res$model_coef
model_step = "Configural" 
item_name = "q4"
x_limits = c(-1,1)
y_limits = c(min(df.large.res$q4), max(df.large.res$q4))
ci_level = .95
model_results = results.invariant$model.configural
lv_name = "lv"

  data_coef = partial.coef
  model_step = "intercepts"
  item_name = "RS6"
  x_limits = c(-1,1)
  y_limits = c(min(DF$RS7), max(DF$RS7))
  model_results = partial.rs.2.picture
  ci_level = .95
  lv_name = "RS"
```

```{r plot-function, include = FALSE}
plot_mgcfa <- function(data_coef, # output from model_coef
                       model_step, # which model
                       item_name, # name of observed item
                       x_limits = c(-1,1), # LV limits to graph
                       y_limits, # Y min and max in data 
                       ci_level, # what ci do you want
                       model_results, # what model results do you want 
                       lv_name # which latent is the observed variable on
                       ){

  # calculate cutoff
  cutoff <- qt(p = (1-ci_level)/2, 
               df = sum(unlist(model_results@Data@nobs)), 
               lower.tail = F)

  # get group variable
  group_var <- model_results@Data@group
  group_labels <- model_results@Data@group.label

  # first get the data
  graph.data <- data_coef %>% # put in tidy coefficients
  filter(model == model_step) %>% # pick a model
  filter(grepl(item_name, term)) %>%  # pick a question
  mutate(group = factor(group, levels = names(table(data_coef$group)), 
                        labels = group_labels))

  # make ribbon data y = slope*x + intercept for ci for slopes 
  ribbondata <- bind_rows(
    data.frame(
    x = seq(from = x_limits[1] - 1,  
            to = x_limits[2] + 1, 
            by = .05), 
    group = unique(graph.data$group)[1]
  ) %>% 
    mutate(ymin = (graph.data %>% filter(op == "=~") %>% 
                 slice_head() %>% pull(estimate) * x) - 
             (cutoff*graph.data %>% filter(op == "=~") %>% 
                    slice_head() %>% pull(std.error)) +
             graph.data %>% filter(op == "~1") %>% 
                 slice_head() %>% pull(estimate), 
           ymax = (graph.data %>% filter(op == "=~") %>% 
                 slice_head() %>% pull(estimate) * x) + 
             (cutoff*graph.data %>% filter(op == "=~") %>% 
                    slice_head() %>% pull(std.error)) +
             graph.data %>% filter(op == "~1") %>% 
                 slice_head() %>% pull(estimate)), 
    data.frame(
      x = seq(from = x_limits[1] - 1,  
              to = x_limits[2] + 1, 
              by = .05), 
      group = unique(graph.data$group)[2]
    ) %>% 
      mutate(ymin = (graph.data %>% filter(op == "=~") %>% 
                   slice_tail() %>% pull(estimate) * x) - 
               (cutoff*graph.data %>% filter(op == "=~") %>% 
                      slice_tail() %>% pull(std.error)) +
               graph.data %>% filter(op == "~1") %>% 
                   slice_tail() %>% pull(estimate), 
             ymax = (graph.data %>% filter(op == "=~") %>% 
                   slice_tail() %>% pull(estimate) * x) + 
               (cutoff*graph.data %>% filter(op == "=~") %>% 
                      slice_tail() %>% pull(std.error)) +
               graph.data %>% filter(op == "~1") %>% 
                   slice_tail() %>% pull(estimate))
  )

  # make point data to draw on the intercepts 
  pointdata <- data.frame(
  x = c(0,0),
  y = graph.data %>% filter(op == "~1") %>% pull(estimate), 
  group = graph.data %>% filter(op == "~1") %>% pull(group),
  ymin = graph.data %>% filter(op == "~1") %>% pull(estimate) - 
    cutoff * graph.data %>% filter(op == "~1") %>% pull(std.error), 
  ymax = graph.data %>% filter(op == "~1") %>% pull(estimate) + 
    cutoff * graph.data %>% filter(op == "~1") %>% pull(std.error)
  )

  # make the line data to draw on the slopes
  linedata <- data.frame(
  slope = graph.data %>% filter(op == "=~") %>% pull(estimate), 
  intercept = graph.data %>% filter(op == "~1") %>% pull(estimate), 
  group = graph.data %>% filter(op == "~1") %>% pull(group)
  )

  # make the distributions for the residuals 
  violindata <- data.frame(
  y = c(rnorm(n = 1000, 
            mean = graph.data %>% filter(op == "~~") %>% 
              slice_head() %>% pull(estimate), 
            sd = graph.data %>% filter(op == "~~") %>% 
              slice_head() %>% pull(std.error)), 
        rnorm(n = 1000, 
            mean = graph.data %>% filter(op == "~~") %>% 
              slice_tail() %>% pull(estimate), 
            sd = graph.data %>% filter(op == "~~") %>% 
              slice_tail() %>% pull(std.error))),
  group = c(rep(graph.data %>% filter(op == "~~") %>% 
              slice_head() %>% pull(group), 1000),
            rep(graph.data %>% filter(op == "~~") %>% 
              slice_tail() %>% pull(group), 1000)), 
  x = 1
  )

  # make the latent mean data for right panel
  latent_means <- lavPredict(model_results, 
                               type = "lv", 
                               label = TRUE,
                               assemble = TRUE, 
                               append.data = TRUE)

  latent_means$lv <- latent_means[ , lv_name]
  latent_means$group <- latent_means[ , group_var]

  # make a plot of the variance
  variance_plot <- 
  ggplot(violindata, aes(x = 1, y = y, color = group, fill = group)) + 
  geom_split_violin() + 
  theme_void() + 
  theme(legend.position = "none") + 
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black")

  # make the plot with intercepts and slopes
  intercept_plot <- 
  ggplot() + 
  # basic set up
  theme_classic() + 
  xlab("Latent Variable") + 
  ylab("Observed Variable") + 
  coord_cartesian(xlim = x_limits, ylim = y_limits) + 
  # plot the intercepts 
  geom_point(data = pointdata, 
             aes(x = x, y = y, color = group), 
             inherit.aes = FALSE) + 
  geom_errorbar(data = pointdata,
                aes(x = x, ymin = ymin, ymax = ymax, color = group), 
                inherit.aes = FALSE, width = .10) + 
  # plot the slopes
  geom_abline(data = linedata,
              aes(slope = slope, intercept = intercept, color = group)) + 
  geom_ribbon(data = ribbondata, 
              aes(x = x, ymin = ymin, ymax = ymax, fill = group), 
              inherit.aes = FALSE, alpha = .2) + 
  scale_color_discrete(name = "Group") + 
  scale_fill_discrete(name = "Group") + 
  geom_vline(xintercept = 0) + 
  theme(axis.line.y = element_blank())

  # make the latent means plot
  mean_plot <- ggplot(latent_means, aes(x = lv, fill = group)) + 
    geom_density(alpha = .2) + 
    theme_classic() + 
    xlab("Latent Variable") + 
    ylab("Density") + 
    geom_vline(data = latent_means %>% group_by(group) %>% summarize(mean = mean(lv)), 
               aes(xintercept = mean, color = group)) + 
    theme(legend.position = "none") + 
    coord_cartesian(xlim = x_limits)

  y_range = abs(y_limits[2] - y_limits[1])

  # line up the two plots 
  prow <- plot_grid(
    intercept_plot + 
      ggtitle("Item Invariance") + 
      theme(legend.position = "none") + 
      annotation_custom(ggplotGrob(variance_plot), 
                        xmin = .25, xmax = 1, 
                        ymin = y_limits[1], ymax = y_limits[2]-y_range/1.8),
    mean_plot + 
      ggtitle("Latent Mean Distribution") + 
      theme(legend.position = "none"),
    align = 'vh',
    hjust = -1,
    nrow = 1
  )

  # get the legend
  legend_b <- get_legend(
    intercept_plot + 
      guides(color = guide_legend(nrow = 1)) +
      theme(legend.position = "bottom")
  )

  # send out the plot 
  plot_grid(prow, legend_b, ncol = 1, rel_heights = c(1, .1))

}
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(84393)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

@@

Measurement invariance is typically analyzed using confirmatory factor analysis, specifically, multi-group confirmatory factor analysis (MGCFA) or less often, with item response theory [@stark2006; @tay2015]. First, the model is examined with the factor structure proposed for the latent and observed variables, and then often these models are assessed for each group separately. The two models are then combined together into one nested CFA in order to determine configural invariance [@brown2015; @byrne2001; @kline2016]. Configural invariance tests if the proposed factor structure is the same between groups. In this model, all other estimated parameters are allowed to vary between groups. The general approach is to use this model as a baseline for starting a sequential analysis of further restrictions between group parameters (i.e., more restrictive with each step). However, models without configural invariance can occur and often point to misspecification for the observed and latent variables within one group (i.e., cross loadings of items onto other latent variables or correlated error terms for one group only).

Next, the estimated parameter between each observed variable and its latent variable are constrained to be equal between groups for metric invariance. For example, item 1's factor loading must be equal to item 1's factor loading for each group. This test examines if the items represent the same relationship to the latent variable, or if specific items have weaker or stronger relationships in specific groups. Finding non-invariance at this stage generally points to items that have different functioning or interpretation for one group. At the third model, the item intercepts (i.e., item averages) are restricted across groups for scalar invariance. Scalar non-invariance would indicate that items have the same strength of relationship with their latent variable, just one group has a higher overall average on that item. Last (although sometimes not used), we may consider constraining error variances for each observed variable to be equal across groups for strict invariance. Strict non-invariance can occur when one group has a higher range of values on the observed variable, thus showing a larger variance. For example, if using a Likert scale, one group may use the full 1 to 7 range (creating a flatter distribution and larger variance), while the other group shows a ceiling effect of only using 5 to 7.

These concepts have been explored and implemented for the last fifty years [@jöreskog1971; @sörbom1978] and implemented in the most popular structural equation modeling programs [@jöreskog2001; @rosseel2012; @boker2011]. @byrne1989 extended the ideas of multi-group testing by suggesting partial invariance [followed by @meredith1993]. Partial invariance occurs when non-invariance is found but can be attributed to only a few parameter estimate differences between groups (i.e., items 1 and 2 have different factor loadings but all others are the same). This testing provided an advantage to understand where the potential non-invariance may occur for further study and interpretation guidelines. To determine when non-invariance and partial invariance occurred, each model is sequentially compared to the previous model using some form of a difference test. Traditionally, since models were nested, a chi-square difference test was used [@meade2008; @cheung2002]; however, given the known issues with chi-square [@thompson1996], people have favored empirical cutoffs for differences in fit indices. As the field pushes back against favoring cutoff criteria and rules of thumb [@marsh2004; @putnick2016], an effect size measure for translating "how much" non-invariance was developed $d_{MACS}$ [@nye2011]. This effect size examines the differences in observed variables between the two groups for both the factor loading and the item intercept; thus, any differences in either or both will increase the effect size for non-invariance [@stark2006].

With $d_{MACS}$ and measurement invariance testing, researchers can begin to quantify how and where their construct measurement may vary between groups. However, given the large number of studies that show non-invariance, it is clear that equivalence can be hard to meet. It is difficult to know if non-invariance occurs because of random sampling error, true population differences, or differences in replication and reproducibility of the construct in a new sample. Further, it is important to remember that the parameter estimates that we are testing are just that - estimates. All the parameter estimates have measures of standard error to indicate that they are more than likely variable with a new sample or population. Given that this information is generally ignored during the examination of measurement invariance, it may be that we are claiming that many scales are non-invariant, when in reality, the differences between loadings or item intercepts are small and unimportant. $d_{MACS}$ provides the opportunity to begin to think about the smallest effect size of interest or the smallest meaningful effect size [@lakens2017; @anvari2021]. As mentioned, $d_{MACS}$ has only really been explored for a combined intercept and loadings, and while useful, does not necessarily allow a researcher to pinpoint specific issues within an observed variable. The purpose of this manuscript is provide readers with a framework for visualization of differences in loadings, intercepts, and variances for each item, and the impact of those differences on the distribution of the latent mean. No known visualization techniques have been proposed for measurement invariance. By creating panel visualizations, we can supplement a researchers ability to judge the strength of the non-invariance differences and effect size for each item. Coupled with other indicators (i.e., fit indices differences, $d_{MACS}$), we can move toward a better understanding of how much measurement non-invariance is meaningful.

@@

1.  Be able to create visualizations for common steps to multi-group confirmatory factor analysis.
2.  Be able to interpret the impact and size of potential non-invariance on measurement.
3.  Understand the impact of measurement variability on replication and generalizability.

@@

Data was simulated using the `simulateData` function in the *R* package *lavaan* [@rosseel2012] assuming multivariate normality using a $\mu$ of 0 and $\sigma$ of 1 for the data. This function allows you to write *lavaan* syntax for your model with estimated values to generate data for observed variables. The data included two groups of individuals ("Group 1", "Group 2") for a multi-group confirmatory factor analysis ($n_{group}$ = 250, *N* = 500). The latent variables were assumed to be continuous normal. The model consisted of five observed items predicted by one latent variable (`lv =~ q1 + q2 + q3 + q4 + q5`); however, the demonstration in this manuscript extends to multiple latent variables and other combinations of observed variables. Each item was assumed to be related to the latent variable with loadings approximately equal to .40 to .80, except when cases of non-invariance on the loadings was assumed.

The @brown2015 steps of testing measurement invariance are demonstrated in this manuscript for illustration purposes, but in line with @stark2006 suggestions, the visualizations show the impact of loadings and intercepts together. The configural model was analyzed nesting both groups into the same CFA model requiring that both groups show the same model structure, but all other parameters are free to vary between groups. The metric model constrained the factor loadings of each group to be equal within the model. The scalar model then constrained the item intercepts (i.e., item means) to be equal across groups. Finally, the strict model constrained the item variances (i.e., error variances) to be equal for each item across groups. These models are normally tested sequentially, and a convenience function `mgcfa` is provided in the supplemental documents for this manuscript. Fit indices for the steps for multi-group models are presented in the appendix for comparison of cutoff rules of thumb [@cheung2002] to effect sizes and visualizations presented in this manuscript. Fit indices include Akaike Information Criterion [AIC, @akaike1998], Bayesian Information Criterion [BIC, @schwarz1978], Comparative Fit Index [CFI, @bentler1990], Tucker Lewis Index [TLI, @tucker1973], root mean squared error of approximation RMSEA [@steiger1990], and standardized root mean square residual [SRMR, @bentler1995].

The data was then simulated to represent invariance across all model steps, small, medium, and large invariance using $d_{MACS}$ estimated sizes from @nye2019. While $d_{MACS}$ is used primarily for an effect size of the (non)-invariance for intercepts and loadings together, a similar approach was taken for the estimation of small, medium, and large effects on the residuals. The effect size is presented for all models, calculated from the *dmacs* package [@dueber2023; @nye2011]. Only one item in each model was manipulated from the invariant model to create the non-invariant models.

# Results

@@

# build invariant model one factor
model.invariant.g1 <- "lv =~ .8*q1 + .4*q2 + .6*q3 + .3*q4 + .6*q5 
q4 ~~ 1*q4
q4 ~ 0*1
q1 ~ 0*1
lv ~ 1"
model.invariant.g2 <- "lv =~ .77*q1 + .43*q2 + .58*q3 + .3*q4 + .61*q5
q4 ~~ 1*q4
q4 ~ 0*1
q1 ~ 0*1
lv ~ 1"

# loadings
model.small.load.g2 <- "lv =~ .77*q1 + .43*q2 + .58*q3 + .5*q4 + .61*q5
q1 ~ 0*1
lv ~ 1"
model.med.load.g2 <- "lv =~ .77*q1 + .43*q2 + .58*q3 + .6*q4 + .61*q5
q1 ~ 0*1
lv ~ 1"
model.large.load.g2 <- "lv =~ .77*q1 + .43*q2 + .58*q3 + 1.1*q4 + .61*q5
q1 ~ 0*1
lv ~ 1"

# build invariance on intercepts .25, .50, and .75
model.small.int.g2 <- "lv =~ .77*q1 + .43*q2 + .58*q3 + .34*q4 + .61*q5
q4 ~ .25*1
q1 ~ 0*1
lv ~ 1"
model.med.int.g2 <- "lv =~ .77*q1 + .43*q2 + .58*q3 + .34*q4 + .61*q5
q4 ~ .5*1
q1 ~ 0*1
lv ~ 1"
model.large.int.g2 <- "lv =~ .77*q1 + .43*q2 + .58*q3 + .34*q4 + .61*q5
q4 ~ .75*1
q1 ~ 0*1
lv ~ 1"

# build invariance on residuals .25, .50, and .75
model.small.res.g2 <- "lv =~ .77*q1 + .43*q2 + .58*q3 + .34*q4 + .61*q5
q4 ~~ .75*q4
q1 ~ 0*1
lv ~ 1"
model.med.res.g2 <- "lv =~ .77*q1 + .43*q2 + .58*q3 + .34*q4 + .61*q5
q4 ~~ .5*q4
q1 ~ 0*1
lv ~ 1"
model.large.res.g2 <- "lv =~ .77*q1 + .43*q2 + .58*q3 + .34*q4 + .61*q5
q4 ~~ .25*q4
q1 ~ 0*1
lv ~ 1"
```

@@

```{r run-cfa}
# invariant CFA
results.invariant <- mgcfa(df.invariant, "group", model.overall)
d_invariant <- lavaan_dmacs(results.invariant$model.configural, RefGroup = "Group 1")

# loadings
results.small.load <- mgcfa(df.small.load, "group", model.overall)
results.med.load <- mgcfa(df.med.load, "group", model.overall)
results.large.load <- mgcfa(df.large.load, "group", model.overall)

d_small_load <- lavaan_dmacs(results.small.load$model.configural, RefGroup = "Group 1")
d_med_load <- lavaan_dmacs(results.med.load$model.configural, RefGroup = "Group 1")
d_large_load <- lavaan_dmacs(results.large.load$model.configural, RefGroup = "Group 1")

# intercepts
results.small.int <- mgcfa(df.small.int, "group", model.overall)
results.med.int <- mgcfa(df.med.int, "group", model.overall)
results.large.int <- mgcfa(df.large.int, "group", model.overall)

d_small_int <- lavaan_dmacs(results.small.int$model.configural, RefGroup = "Group 1")
d_med_int <- lavaan_dmacs(results.med.int$model.configural, RefGroup = "Group 1")
d_large_int <- lavaan_dmacs(results.large.int$model.configural, RefGroup = "Group 1")

# residuals
results.small.res <- mgcfa(df.small.res, "group", model.overall)
results.med.res <- mgcfa(df.med.res, "group", model.overall)
results.large.res <- mgcfa(df.large.res, "group", model.overall)

>>>>

---
title             : "visualizemi: Visualization, Effect Size, and Replication of Measurement Invariance for Registered Reports"
shorttitle        : "VISUAL MGCFA"

@@

abstract: |
  Latent variable modeling as a lens for psychometric theory is a popular tool for social scientists to examine measurement of constructs [@beaujean2014]. Journals such as *Assessment* regularly publish articles supporting measures of latent constructs wherein a measurement model is established. Confirmatory factor analysis can be used to investigate the replicability and generalizability of the measurement model in new samples, while multi-group confirmatory factor analysis is used to examine the measurement model across groups within samples [@brown2015]. With the rise of the replication crisis and "psychology's renaissance" [@nelson2018], interest in divergence in measurement has increased, often focused on small parameter differences within the latent model. This manuscript presents `visualizemi`, an *R* package that provides functionality to calculate multigroup models, partial invariance, visualizations for (non)-invariance, effect sizes for models and parameters, and potential replication rates compared to random models. Readers will learn how to interpret the impact and size of the proposed non-invariance in models with a focus on potential replication and how to plan for registered reports.

@@

classoption       : "man"
bibliography: includes/references.bib

@@

csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa7"
output            : papaja::apa6_pdf
appendix          :
  - "appendix_1.Rmd"
  - "appendix_2.Rmd"
  - "appendix_3.Rmd"
  - "appendix_4.Rmd"
  - "appendix_5.Rmd"
  - "appendix_6.Rmd"
---

```{r libraries, include = FALSE}
library(papaja)
library(lavaan)
library(broom)
library(dplyr)
library(ggplot2)
library(cowplot)
library(ggridges)
library(tidyr)
library(rio)

# devtools::install_github("psyteachr/introdataviz")
# devtools::install_github("ddueber/dmacs")
# devtools::install_github("doomlab/visualizemi")
library(introdataviz)
library(dmacs)
library(visualizemi)

@@

```{r analysis-preferences}
# Seed for random number generation
set.seed(84393)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, cache = TRUE, message = FALSE, warning = FALSE)

boot_rr_DF <- import("manu_data/boot_rr_DF.csv")
boot.partial.invariant <- readRDS("manu_data/boot.partial.invariant.RData")
boot.model.rs <- readRDS("manu_data/boot.model.rs.RData")
boot.partial.rs <- readRDS("manu_data/boot.partial.rs.RData")

boot.df.full <- import("manu_data/boot_DF_full.csv")
boot.effects.summary.full <- import("manu_data/boot_effects_summary_DF_full.csv")
boot.partial.summary.full <- import("manu_data/boot_partial_summary_DF_full.csv")
```

@@

Measurement invariance has been explored and implemented for the last fifty years [@jöreskog1971; @sörbom1978] and implemented in the most popular structural equation modeling programs [@jöreskog2001; @rosseel2012; @boker2011]. @byrne1989 extended the ideas of multi-group testing by suggesting partial invariance [followed by @meredith1993]. Partial invariance occurs when non-invariance is found but can be attributed to only a few parameter estimate differences between groups (i.e., items 1 and 2 have different factor loadings but all others are the same). This testing provided an advantage to understand where the potential non-invariance may occur for further study and interpretation guidelines. As the field pushes back against favoring cutoff criteria and rules of thumb [@marsh2004; @putnick2016], an effect size measure for translating "how much" non-invariance was developed $d_{MACS}$ [@nye2011]. This effect size examines the differences in observed variables between the two groups for both the factor loading and the item intercept; thus, any differences in either or both will increase the effect size for non-invariance [@stark2006].

With $d_{MACS}$ and measurement invariance testing, researchers can begin to quantify how and where their construct measurement may vary between groups. Yet, given the large number of studies that show non-invariance, it is clear that equivalence can be hard to meet. It is difficult to know if non-invariance occurs because of random sampling error, true population differences, or differences in replication and reproducibility of the construct in a new sample. The field of psychology is increasingly interested in pre-registration [i.e., registering plans for a study before data collection, @nosek2018] and the promotion of transparency in study design, implementation, and analysis [@mayo-wilson2021], in addition to supporting replication studies [@zwaan2018]. Registered (replication) reports provide an advantageous avenue for the pre-registration of measurement tests, as they allow a researcher the ability to have their study accepted in principle, regardless of the results of a test of construct validity, reliability, or measurement invariance [@hobson2019; @nosek2014]. However, there are few tools that can provide effect size measures for models, individual parameters, or visualization for researchers to plan for future studies. $d_{MACS}$ provides the opportunity to begin to think about the smallest effect size of interest or the smallest meaningful effect size for measurement invariance and replication [i.e., two studies with overlapping confidence intervals "replicate", even if the test of measurement invariance does not, @lakens2017; @anvari2021]. As mentioned, $d_{MACS}$ has only really been explored for a combined intercept and loadings, and while useful, does not necessarily allow a researcher to pinpoint specific issues within an observed variable.

Therefore, purpose of this manuscript is describe an *R* package, `visualizemi`, that provides functionality to calculate multi-group confirmatory factor analysis, partial invariance tests, visualizations of the size of non-invariance, and potential effect sizes for overall models and individual parameters. No known visualization techniques have been proposed for measurement invariance. By creating panel visualizations, we can supplement a researchers ability to judge the strength of the non-invariance differences and effect size for each item. The proposed effect sizes demonstrate the likelihood of replication with a similar sample as compared to a randomly assigned group model, thus, illustrating what type of measurement one might expect to find, and how different that is from random chance. Within this technique, the individual parameter effect sizes can calculated: both the group differences within a model as compared to random and the likelihood of a parameter replication compared to random groups. Coupled with other indicators (i.e., fit indices differences, $d_{MACS}$), we can move toward a better understanding of how much measurement non-invariance is meaningful. This tutorial and package will help researchers plan future studies and aid in the ability to estimate a smallest effect of interest for measurement invariance studies, rather than relying on fit indices and rules of thumb alone.

@@

1.  Learn how to use *visualizemi* to analyze multi-group confirmatory factor analysis, examine partial invariance, and create visualizations of parameters.
2.  Learn how to estimate the potential replication of multi-group models and their parameters using bootstrapping compared to a random group model.
3.  Be able to calculate and interpret effect sizes for model and parameter replication, as well as parameter group differences.
4.  Understand the impact of measurement variability on replication and generalizability.

The tutorial will start with simulated data based on known effect sizes using $d_{MACS}$ and demonstrating the package functions for 1) running the multigroup analysis, 2) running a partial invariance analysis, 3) plotting the partial invariance, 4) estimating replication and effect sizes at the model level, and 5) estimating replication and effect sizes at the parameter level. Last, data from @aiena2014 examining the measurement invariance of the RS-14 [@wagnild2009] will be used to demonstrate the application of the package on real data. The *visualizemi* package vignette includes an additional tutorial walk through.

@@

Data was simulated using the `simulateData` function in the *R* package *lavaan* [@rosseel2012] assuming multivariate normality using a $\mu$ of 0 and $\sigma$ of 1 for the data. This function allows you to write *lavaan* syntax for your model with estimated values to generate data for observed variables (see supplemental for examples). The data included two groups of individuals ("Group 1", "Group 2") for a multi-group confirmatory factor analysis ($n_{group}$ = 250, *N* = 500). The latent variables were assumed to be continuous normal (the package functions do not require this assumption). The model consisted of five observed items predicted by one latent variable (`lv =~ q1 + q2 + q3 + q4 + q5`); however, the demonstration in this manuscript extends to multiple latent variables and other combinations of observed variables. Each item was assumed to be related to the latent variable with loadings approximately equal to .40 to .80, except when cases of non-invariance on the loadings was simulated.

The @brown2015 steps of testing measurement invariance are demonstrated in this manuscript for illustration purposes, but in line with @stark2006 suggestions, the visualizations show the impact of loadings and intercepts together. A convenience function `mgcfa` is used for these steps or other measurement invariance test orders and combinations. Fit indices for the steps for multi-group models are presented in the appendix for comparison of cutoff rules of thumb [@cheung2002] to effect sizes and visualizations presented in this manuscript. Fit indices include Akaike Information Criterion [AIC, @akaike1998], Bayesian Information Criterion [BIC, @schwarz1978], Comparative Fit Index [CFI, @bentler1990], Tucker Lewis Index [TLI, @tucker1973], root mean squared error of approximation RMSEA [@steiger1990], and standardized root mean square residual [SRMR, @bentler1995].

The data was then simulated to represent invariance across all model steps, small, medium, and large invariance using $d_{MACS}$ estimated sizes from @nye2019. While $d_{MACS}$ is used primarily for an effect size of the (non)-invariance for intercepts and loadings together, a similar approach was taken for the estimation of small, medium, and large effects on the residuals. The effect size is presented for all models, calculated from the *dmacs* package [@dueber2023; @nye2011]. Only one item in each model was manipulated from the invariant model to create the non-invariant models. Given the data was simulated with a *z*-score scaling, the loading values were simulated at .30 points apart (given $d_{MACS}$ suggestions of .2, .4, .7), the intercepts at .25 points apart, and the residuals at .25 points apart. To plan a simulation for your own study, these values can be used to simulate small, medium, and large non-invariance effects by first converting data into *z*-score.

# Package Code Examples

The complete code for this manuscript can be found at <https://osf.io/wev5f/>. This tutorial was registered at <https://osf.io/vwf4d>, and the example provided at the end of the manuscript was added after that registration. The *R* package and replication/effect sizes was added after the original manuscript submission.

@@

# build invariant model one factor
model.invariant.g1 <- "lv =~ .8*q1 + .4*q2 + .6*q3 + .3*q4 + .6*q5 
q4 ~~ 1*q4
q4 ~ 0*1
q1 ~ 0*1
lv ~ 1"
model.invariant.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .31*q4 + .61*q5
q4 ~~ 1*q4
q4 ~ 0*1
q1 ~ 0*1
lv ~ 1"

# loadings .5, .8, 1.1
model.small.load.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .5*q4 + .61*q5
q1 ~ 0*1
lv ~ 1"
model.med.load.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .8*q4 + .61*q5
q1 ~ 0*1
lv ~ 1"
model.large.load.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + 1.1*q4 + .61*q5
q1 ~ 0*1
lv ~ 1"

# build invariance on intercepts .25, .50, and .75
model.small.int.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .31*q4 + .61*q5
q4 ~ .25*1
q1 ~ 0*1
lv ~ 1"
model.med.int.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .31*q4 + .61*q5
q4 ~ .5*1
q1 ~ 0*1
lv ~ 1"
model.large.int.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .31*q4 + .61*q5
q4 ~ .75*1
q1 ~ 0*1
lv ~ 1"

# build invariance on residuals .25, .50, and .75
model.small.res.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .31*q4 + .61*q5
q4 ~~ .75*q4
q1 ~ 0*1
lv ~ 1"
model.med.res.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .31*q4 + .61*q5
q4 ~~ .5*q4
q1 ~ 0*1
lv ~ 1"
model.large.res.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .31*q4 + .61*q5
q4 ~~ .25*q4
q1 ~ 0*1
lv ~ 1"
```

@@

```{r run-cfa}
# invariant CFA
results.invariant <- 
  mgcfa(model = model.overall, 
        data = df.invariant,
        group = "group",
        group.equal = c("loadings", "intercepts", "residuals"),
        meanstructure = T)

d_invariant <- lavaan_dmacs(
  results.invariant$model_configural, 
  RefGroup = "Group 1")

# loadings
results.small.load <- 
  mgcfa(model = model.overall, 
        data = df.small.load,
        group = "group",
        group.equal = c("loadings", "intercepts", "residuals"),
        meanstructure = T)

results.med.load <- 
  mgcfa(model = model.overall, 
        data = df.med.load,
        group = "group",
        group.equal = c("loadings", "intercepts", "residuals"),
        meanstructure = T)
results.large.load <- 
  mgcfa(model = model.overall, 
        data = df.large.load,
        group = "group",
        group.equal = c("loadings", "intercepts", "residuals"),
        meanstructure = T)

d_small_load <- 
  lavaan_dmacs(results.small.load$model_configural, 
               RefGroup = "Group 1")
d_med_load <- 
  lavaan_dmacs(results.med.load$model_configural, 
               RefGroup = "Group 1")
d_large_load <- 
  lavaan_dmacs(results.large.load$model_configural, 
               RefGroup = "Group 1")
